--- 
title: "Third and Six"
author: "Ben Reitman"
date: '2023-10-17'
format:
  html:
    code-fold: true
---

```{r}
library(tidyverse)
library(cfbfastR)
```

# load play by play data
```{r}
plays <- load_cfb_pbp(seasons = c(2019, 2020, 2021, 2022, 2023))
```



```{r}
third_and_six <- plays %>% 
  select(year,pos_team, down, distance, play_type, yards_gained) %>% 
  filter(down == 3, distance == 6)
  
```


```{r}
play_type_count<-third_and_six %>% 
  group_by(play_type) %>% 
  summarise(count = n())


```
## Of all of the plays run on 3rd and 6, 360 of them were rushing plays. In addition, 20 plays run on 3rd and 6 ended up in rushing touchdowns.


```{r}
play_type_count %>% 
  mutate(total = sum(count)) %>% 
  mutate(percent = (count/total) *100) %>% 
  arrange(desc(percent))
```
## 18% of plays run on third and six were runs and an additional 1% of rush plays run on third and six resulted in touchdowns.




```{r}
rushing_plays <- third_and_six %>% 
  filter(down == 3, distance == 6, play_type == "Rush") %>% 
  mutate(mean(yards_gained))
```
## We see that the mean yard average of runs on third down and six on plays that did not end in touchdowns is 5.86. 


```{r}
rushing2019<-rushing_plays %>%  
  filter(year == '2019')

rushing2020<-rushing_plays %>%  
  filter(year == '2020')

rushing2021<-rushing_plays %>%  
  filter(year == '2021')

rushing2022<-rushing_plays %>%  
  filter(year == '2022')

rushing2023<-rushing_plays %>%  
  filter(year == '2023')
```

## Out of the four most recent completed seasons and this season so far, teams ran the ball on 3rd and 6 the most in 2022, 748 times. This is more than 200 than the next closest season which is 2021. 


```{r}
rushing2019 %>% 
  group_by(pos_team) %>% 
  summarise(count =n()) %>% 
  arrange(desc(count))


rushing2020 %>% 
  group_by(pos_team) %>% 
  summarise(count =n()) %>% 
  arrange(desc(count))

rushing2021 %>% 
  group_by(pos_team) %>% 
  summarise(count =n()) %>% 
  arrange(desc(count))

rushing2022 %>% 
  group_by(pos_team) %>% 
  summarise(count =n()) %>% 
  arrange(desc(count))

rushing2023 %>% 
  group_by(pos_team) %>% 
  summarise(count =n()) %>% 
  arrange(desc(count))
```
## So far this season, South Florida has run the ball on 3rd and 6 the most doing it 6 times. In the other four recent years, Navy led the nation twice, Army did it once, and Georgia Southern did it once. Navy and Army both run the triple option so it makes a lot of sense that they would be at or near the top of these lists. 


```{r}
Maryland <- third_and_six %>% 
  filter(pos_team == "Maryland")
```


```{r}
Maryland %>% 
  group_by(play_type) %>% 
  summarise(count=n()) %>% 
  arrange(desc(count))

```
## It seems that the third and six run by Maryland on Saturday against Illinois was not an anamoly. Maryland has rushed the ball 15 times on third and six in the last five seasons. They have thrown the ball 24 times, completing 13 of those passes. Maryland has also rushed for one touchdown on 3rd and 6 in the last five years as well. 



